---
title: "Multivariate MCA"
author: "Susan Holmes"
date: "April 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load the data : supposing it is in the data directory


This is how we read in the data:
(your file will be in a different place)
```{r}
CP=read.csv("./data/CP1_Alla.csv",stringsAsFactors = TRUE,header=TRUE)
```

```{r}
CPb=CP
CPb$LT=as.factor(CP$LT)
CP.tab=xtabs(Frequency~Lamin+Chromatin+LT+Type,data=CPb)
CP.tab=xtabs(Frequency~Lamin+Chromatin+LT+Type,data=CPb)
freqdata=data.frame(CP.tab)
```
A little testing using a glm for counts (rare)
```{r}
model=glm(Freq~Type,family=poisson,data=freqdata)
anova(model,test="Chisq")
fullmodel=glm(Freq~Type+Lamin+Chromatin,family=poisson,data=freqdata)
anova(fullmodel,test="Chisq")
```

Not much difference between the two types, but significance for Lamin
and Chromatin.



```{r}
attach(freqdata)
xtabA <- tapply(Freq,list(Type,Lamin,Chromatin,LT),c)
detach(freqdata)
ftab=ftable(xtabA)
ftab
```

You can do a whole set of independence tests by doing:
```{r}
xtabA[1,,,]
mantelhaen.test(xtabA[1,,,])
xtabA[2,,,]
mantelhaen.test(xtabA[2,,,])
```

We see the top one is significant the bottom one less so,
you can do the others in the same way, but this does not show all the
dimensions of variation.

You'll need some new packages to install:
```{r, eval=FALSE}
install.packages(c("ade4","factoextra","reshape","FactoMineR","vcd"))
```

This plot shows the cells in the table that are not 
as they would have been under independence, red low, blue high.
```{r mosaicplot}
library(vcd)
mosaic(CP.tab,shade=TRUE)
```
It's hard to read, so we're going to do multifactorial analyses
(a sort of PCA for contingency tables)

```{r}
library(ade4)
library(reshape)
library(FactoMineR)
library(factoextra)
disj=tab.disjonctif(freqdata[,-5])
expanded=untable(disj,num=freqdata[,5])
##Need different column names for all categories
colnames(expanded)<-c("Lam_Abs","Lam_Cont","Lam_Deg","Chrom_Abs","Chrom_Base","Chrom_Deg","LT0","LT1","CP1","Lucif")
burtdf=as.data.frame(expanded)
rownames(burtdf)=paste("r",1:454,sep="")
```

```{r mca1}
bbcoa <- dudi.coa(burtdf,scannf=FALSE,nf=4)
s.arrow(bbcoa$c1,boxes=FALSE)
s.arrow(bbcoa$c1[,3:4],boxes=FALSE)
```



```{r fancierbiplots}
fviz_ca_biplot(bbcoa, repel = FALSE)
fviz_ca_biplot(bbcoa, repel = TRUE)
fviz_ca_biplot(bbcoa, repel = FALSE,axes=c(3,4))
```

## New idea : separate Luciferase and CP1

```{r lucif}
Lucifonly=burtdf[burtdf[,10]==1,-c(9,10)]
Lucoa <- dudi.coa(Lucifonly,scannf=FALSE,nf=4)
s.arrow(Lucoa$c1,boxes=FALSE)
```

```{r CP1}
CP1only=burtdf[burtdf[,9]==1,-c(9,10)]
CP1coa <- dudi.coa(CP1only,scannf=FALSE,nf=4)
s.arrow(CP1coa$c1,boxes=FALSE)
```

```{r sidebyside}
par(mfrow=c(1,2))
s.arrow(Lucoa$c1,boxes=FALSE)
s.arrow(CP1coa$c1,boxes=FALSE)
par(mfrow=c(1,1))
```
So not to be confused by a simple change in sign on the second
axis (which is meaningless),we switch those coefficients.

```{r}
CP1coa$c1[,2]=-CP1coa$c1[,2]
par(mfrow=c(1,2))
s.arrow(Lucoa$c1,boxes=FALSE)
s.arrow(CP1coa$c1,boxes=FALSE)
par(mfrow=c(1,1))

```

There seems to be unfortunately very little difference in what is going on in the two dynamics, except things are more gradual in the CP1
case as more of the states "are intermediary".

Not quite sure about the inversion of Chrom_Abs and Lam_Abs.


